Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality

### 2015-07-14 content/contribution/CoDeR.tex
Cleaning up §3, minor changes elsewhere.

\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^C = \langle R,F,S,r \rangle \vDash \langle E,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
	It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.

@@

		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^C$ expressed by a CoDeR-based system.
		These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
		an instance of an RDF graph is an RDF graph annotated with a time interval: ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the input stream must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the time interval of the graph instance in which it originated.

		The interval ${(i_e,i_n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^C = \langle R, F, S, r \rangle}$ expressed.
		For graphs on the input stream, $i_e$ is the time at which a given graph was asserted, and ${i_n = i_e + r}$ is the time at which it will leave the semantic sliding window over $S$ of $\Pi^C$.
		For graphs on the output stream, $i_e$ is the latest time at which any part of the justification of a given graph was asserted, and $i_n$ is the earliest time at which any part of its justification will leave the sliding window.
		\reffig{intersected-intervals} provides an example of this for some ${\Pi^C = \langle \{c\,\text{:-}\,a,b\,.\},\{\},(\{a\}|1,\{b\}|2),3\}}$ over instants ${i = 1 \dots 4}$.
		\begin{figure}
			\centering
			\includegraphics[width=0.45\textwidth]{intersected-intervals}
			\caption{Derivation of valid-time for entailed fact instance from those of its justifying fact instances.}
			\labelfig{intersected-intervals}
		\end{figure}
		\begin{align*}
			\Pi^C_{1} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \} \rangle & \vDash & \{ \} \\
			\Pi^C_{2} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \, b|2. \} \rangle & \vDash & \{ c|2. \} \\
			\Pi^C_{3} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \, b|2. \} \rangle & \vDash & \{ c|2. \} \\
			\Pi^C_{4} & = \langle \{c\,\text{:-}\,a,b.\}, \{ b|2. \} & \vDash & \{ \}
		\end{align*}
		It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^C$.
		However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^C$ ordered by the instant $i_{n}$ at which they \emph{ceased} to be justified by $W^{S,r}_{i_n}$, the output stream of a CoDeR-based system also expresses the stream $N$.
		As such, the annotation of streamed RDF graphs with the interval $(i_e,i_n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.

		In the second case, inter-operator streams are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
		These streams are sequences of the intermediate results of processing between steps in the production of entailments.
		As the production of entailments in CoDeR is cast as a pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
		an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^C$.
		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
		It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
		the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are either the entailments of the system or the assertions of its predecessor in the stream pipeline.

@@

		As all operators are stream-to-stream in nature, they each take one or two streams as input and produce a single stream (though they may contribute to any number of union streams).
		As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be used as input to another operator.
		For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${\Pi^C = \{ \{b\,\text{:-}\,a.\}, \{\}, S, r \}}$, where $a$ and $b$ are RDF triples.
		\begin{equation}\labeleqn{basic-example-plan}
			\text{:-}_b \left( \sigma_a \left( {}_g{\pi_t}\,S \right) \right)
		\end{equation}

@@

				Instance implication is a one-pass operator that takes as input an inter-operator stream $X$ whose graph pattern is the body of one of the rules $R$ of the expressed $\Pi^C$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.

@@

		% As the SteM ${{}_b{\Join_p}}$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
		As the SteM $\rstreamjoin$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
		Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
		as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $r$ over the system input streams $S$ defined in a given ${\Pi^C = \{R,F,S,r\}}$.
		It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of Continuous Datalog.
		It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^C$ at a given instant will be persisted within that system at that instant,
		i.e. such a CoDeR-based system is guaranteed to produce the complete set of results according to the semantics of Continuous Datalog.
	\end{nestedsection}
\end{nestedsection}

>>>>

\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^{C,r} = \langle R,F^0,S,r \rangle \models \langle E^0,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
	It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.

@@

		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^{C,r}$ expressed by a CoDeR-based system.
		These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
		an instance of an RDF graph is an RDF graph annotated with a time interval: ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the input stream must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the time interval of the graph instance in which it originated.

		The interval ${(i_e,i_n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ expressed.
		For graphs on the input stream, $i_e$ is the time at which a given graph was asserted, and ${i_n = i_e + r}$ is the time at which it will leave the semantic sliding window over $S$ of $\Pi^{C,r}$.
		For graphs on the output stream, $i_e$ is the latest time at which any part of the justification of a given graph was asserted, and $i_n$ is the earliest time at which any part of its justification will leave the sliding window.
		\reffig{intersected-intervals} provides an example of this for some ${\Pi^{C,r} = \langle \{c \Leftarrow a \land b\},\{\},[\langle a,1 \rangle,\langle b,2 \rangle ], 3\rangle}$ over instants ${i = 1 \dots 4}$.
		\begin{figure}
			\centering
			\includegraphics[width=0.25\textwidth]{intersected-intervals}
			\caption{Derivation of valid-time for entailed fact instance from those of its justifying fact instances.}
			\labelfig{intersected-intervals}
		\end{figure}
		\begin{align*}
			\Pi^{C,r}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
			\Pi^{C,r}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
			\Pi^{C,r}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
			\Pi^{C,r}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
		\end{align*}
		It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^{C,r}$.
		However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^{C,r}$ ordered by the instant $i_{n}$ at which they \emph{ceased} to be justified by $W^{S,r}_{i_n}$, the output stream of a CoDeR-based system also expresses the stream $N$.
		As such, the annotation of streamed RDF graphs with the interval $(i_e,i_n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.

		In the second case, inter-operator streams are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
		These streams are sequences of the intermediate results of processing between steps in the production of entailments.
		As the production of entailments in CoDeR is cast as a pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
		an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^{C,r}$.
		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
		It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
		the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are either the entailments of the system or the assertions of its predecessor in the stream pipeline.

@@

		As all operators are stream-to-stream in nature, they each take one or two streams as input and produce a single stream (though they may contribute to any number of union streams).
		As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be used as input to another operator.
		For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${\Pi^{C,r} = \langle \{b \Leftarrow a\}, \{\}, S, r \rangle}$, where $a$ and $b$ are RDF triples.
		\begin{equation}\labeleqn{basic-example-plan}
			\text{:-}_b \left( \sigma_a \left( {}_g{\pi_t}\,S \right) \right)
		\end{equation}

@@

				Instance implication is a one-pass operator that takes as input an inter-operator stream $X$ whose graph pattern is the body of one of the rules $R$ of the expressed $\Pi^{C,r}$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.

@@

		% As the SteM ${{}_b{\Join_p}}$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
		As the SteM $\rstreamjoin$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
		Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
		as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $r$ over the system input streams $S$ defined in a given ${\Pi^{C,r} = \langle R,F^0,S,r\rangle}$.
		It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of Continuous Datalog.
		It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^{C,r}$ at a given instant will be persisted within that system at that instant,
		i.e. such a CoDeR-based system is guaranteed to produce the complete set of results according to the semantics of Continuous Datalog.
	\end{nestedsection}
\end{nestedsection}

### 2015-07-14 content/evaluation.tex
Cleaning up §3, minor changes elsewhere.

		Before we move on to more detailed analysis of the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
		However, when we changed the `And' into a series of `seq's, the system runs 20x times faster (7.5 minutes for the same test).
		Such a drastic difference is likely to be caused by the optimisation of Etalis for the \emph{seq} operator: being intended for \emph{event processing} rather than continuous querying/reasoning, the order of arrival of triples is often more relevant than their simple coincidence in the system.
		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating ETALIS to use \emph{seq} instead of \emph{And}.
		It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time- and memory-efficiency of the three systems.
		\begin{figure}
			\centering
			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
			\caption{A comparison of the times taken by the systems R4, Sparkwave and ETALIS to process 1.1 million triples from a single input stream, given sliding windows of various ranges over the input stream.}
			\labelfig{all-systems-varying-windows}
		\end{figure}

@@

		We also examined the memory consumption of R4, Sparkwave, and Jtalis throughout a run.
		Setting the sliding window range to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
		It shows that, while R4 and Jtalis follow similar patterns, Sparkwave memory needs are significantly higher.
		Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to the manual garbage collection performed.
		\begin{figure}
			\centering
			\includegraphics[width=0.5\textwidth]{memoryConsumptionComparison}
			\caption{A comparison of the memory consumed by the systems R4, Sparkwave and ETALIS throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
			\labelfig{all-systems-memory-consumption}
		\end{figure}
	\end{nestedsection}

>>>>

		Before we move on to more detailed analysis of the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
		However, when we changed the `And' into a series of `seq's, the system runs 20x times faster (7.5 minutes for the same test).
		Such a drastic difference is likely to be caused by the optimisation of Etalis for the \emph{seq} operator: being intended for \emph{event processing} rather than continuous querying/reasoning, the order of arrival of triples is often more relevant than their simple coincidence in the system.
		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating Etalis to use \emph{seq} instead of \emph{And}.
		It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time- and memory-efficiency of the three systems.
		\begin{figure}
			\centering
			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
			\caption{A comparison of the times taken by the systems R4, Sparkwave and Etalis to process 1.1 million triples from a single input stream, given sliding windows of various ranges over the input stream.}
			\labelfig{all-systems-varying-windows}
		\end{figure}

@@

		We also examined the memory consumption of R4, Sparkwave, and Jtalis throughout a run.
		Setting the sliding window range to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
		It shows that, while R4 and Jtalis follow similar patterns, Sparkwave memory needs are significantly higher.
		Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to the manual garbage collection performed.
		\begin{figure}
			\centering
			\includegraphics[width=0.5\textwidth]{memoryConsumptionComparison}
			\caption{A comparison of the memory consumed by the systems R4, Sparkwave and Etalis throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
			\labelfig{all-systems-memory-consumption}
		\end{figure}
	\end{nestedsection}

### 2015-07-14 content/personal-details-ieee.tex
Added the CORRECT article class, and set up the template correctly apart from the problem listed below.  Also fixed the details for Rehab in all the personal-details-*.tex
PROBLEMS
- inductive step still overflows both column and page
- abstract or the rest of the frontmatter does not format properly in twocolumn mode
Removed middle initial
Cleaning up §3, minor changes elsewhere.

% Personal Details
\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
\begin{IEEEkeywords}
      Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation
\end{IEEEkeywords}
%\keywords   {Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation}
\author    {\IEEEauthorblockN{David L. Monks, Rehab ???, Dr. Nicholas M. Gibbins}
            \IEEEauthorblockA{
                              Web and Internet Science\\
                              Electronics and Computer Science\\
                              Physical Sciences and Engineering\\
                              University of Southampton\\
                              \{dm11g08,r???,nmg\}@ecs.soton.ac.uk
                              }
            }
% TODO fix Rehab's details
\date{\today}

>>>>

% Personal Details
\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
\begin{IEEEkeywords}
      Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation
\end{IEEEkeywords}
%\keywords   {Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation}
\author    {\IEEEauthorblockN{David L. Monks, Rehab Albeladi, Nicholas Gibbins}
            \IEEEauthorblockA{
                              Electronics and Computer Science\\
                              University of Southampton\\       
                              Southampton, SO17 1BJ\\
                              United Kingdom \\
                              \{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk
                              }
            }
% TODO fix Rehab's details
\date{\today}

\maketitle

\begin{abstract}
	\input{abstract.tex}
\end{abstract}

### 2015-07-14 content/personal-details-ios.tex
Cleaning up §3, minor changes elsewhere.

	\author{\fnms{David L.} \snm{Monks}}
	\sep
	\author{\fnms{Rehab} \snm{Albeladi}}
	and
	\author{\fnms{Nicholas} \snm{Gibbins}}
	\runningauthor{D.L. Monks et al.}
	\address{
	          Department of Electronics and Computer Science,\\
	          Faculty of Physical Sciences and Engineering,\\
	          University of Southampton,\\
	          University Rd., Southampton, SO17 1BJ,\\
	          United Kingdom\\
	          E-mail: \{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk
	}
	% TODO fix Rehab's details

>>>>

	\author{\fnms{David L.} \snm{Monks}}
	\sep
	\author{\fnms{Rehab} \snm{Albeladi}}
	and
	\author{\fnms{Nicholas} \snm{Gibbins}}
	\runningauthor{D.L. Monks et al.}
	\address{
	          Electronics and Computer Science,\\
	          University of Southampton,\\
	          Southampton, SO17 1BJ,\\
	          United Kingdom\\
	          E-mail: \{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk
	}
	% TODO fix Rehab's details
